---
title: "HW6"
author: "Duzhi Zhao"
date: "11/18/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8, 
  fig.height = 6,
  out.width = "100%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1
```{r}
# Import and tidy dataset
birth_w = read.csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  # Change variable type from numeric to ordered factor
  mutate(
    babysex = ordered(babysex, levels = c(1, 2)),
    frace = ordered(frace, levels = c(1, 2, 3, 4, 8)),
    malform = ordered(malform, levels = c(0, 1)),
    mrace = ordered(mrace, levels = c(1, 2, 3, 4))
  ) %>% 
  # Recode variables with direct information
  mutate(
    babysex = recode(babysex, 
                     `1` = "male",
                     `2` = "female"),
    frace = recode(frace, 
                   `1` = "White",
                   `2` = "Black",
                   `3` = "Asian",
                   `4` = "Puerto Rican",
                   `8` = "Other"),
    malform = recode(malform, 
                     `0` = "absent",
                     `1` = "present"),
    mrace = recode(mrace,
                   `1` = "White",
                   `2` = "Black",
                   `3` = "Asian",
                   `4` = "Puerto Rican")
  )

# Check missing value
summary(is.na(birth_w)) # all have FALSE: 4342 --> NO MISSING VALUE
```

```{r}
# Check normality of bwt data
birth_w %>% 
  ggplot(aes(x = bwt)) +
  geom_histogram(bins = 50) +
  labs(x = "Birthweight (g)",
       y = "Count",
       title = "Normality Check on birthweight")
```

**Comments:**  
Birthweight data is approximately following a normal distribution. No transformation is needed in this case.

```{r}
# Multiple linear regression: 
# Predictor var 1: average number of cigarettes smoked per day during pregnancy
# Predictor var 2: mother's age at delivery
# Outcome var: birthweight

fit1 = lm(bwt ~ smoken + momage, data = birth_w)
summary(fit1)

fit1 %>% 
  broom::tidy() %>% 
  mutate(term = str_replace(term, "^smoken", "Average number of cigarettes smoked per day")) %>%
  mutate(term = str_replace(term, "^momage", "Mother's age at delivery")) %>% 
  knitr::kable(digits = 3)
```

**Comments:**  
The multiple linear regression mode that I am interested in is how the average number of cigarettes smoked by the pregnant women per day ('smoken') and their age at delivery ('momage') predict the baby's birth weight ('bwt').  
Results show that for every unit increase in the average number of cigarettes smoked per day during pregnancy, baby's birthweight decreases by 5.971 g, adjusting for mother's age at delivery. For every unit increase in mother's age at delivery, baby's birthweight increases by 18.792 g, adjusting for average number of cigarettes smoked per day during pregnancy. The low adjusted R-squared (0.0254) shows that only 2.54 % of the variation in the outcome variable can be explained by the predictor variables, adjusting for the number of predictors.

```{r}
# Add residuals and predictions (fitted values) into dataset 'birth_w'
bwt_resid_pred = 
  birth_w %>% 
  add_residuals(fit1) %>% 
  add_predictions(fit1)

# Plot: Residual vs. Fitted value 
bwt_resid_pred %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.4) + 
  geom_abline(intercept = 0, slope = 0, color = "red") +
  labs(x = "Predicted/Fitted value",
      y = "Residual",
      title = "Residual vs. Fitted value")
```

**Comments:**  
The residual vs. Fitted value plot shows a random pattern with evenly distributed value around the zero line, indicating equal error variance.

```{r}
# MLR
# Predictor 1: baby's length at birth
# Predictor 2: gestational age in weeks
# Outcome: birthweight

fit2 = lm(bwt ~ blength + gaweeks, data = birth_w)
summary(fit2)
```

```{r}
# MLR - three-way interaction effect
# Predictor 1: baby’s head circumference at birth (centimeters)
# Predictor 2: baby's length at birth
# Predictor 3: baby's sex
# Outcome: birthweight

fit3 = lm(bwt ~ bhead * blength * babysex, data = birth_w)
summary(fit3)
```

```{r}
set.seed(20)

# Cross validating
cv_birthw = crossv_mc(birth_w, 100)
```

```{r}
# Fitting models
cv_results =
  cv_birthw %>% 
  mutate(
    model1 = map(.x = train, ~lm(bwt ~ smoken + momage, data = .x)),
    rmse_model1 = map2_dbl(.x = model1, .y = test, ~rmse(.x, .y)),
    
    model2 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    rmse_model2 = map2_dbl(.x = model2, .y = test, ~rmse(.x, .y)),
    
    model3  = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x)),
    rmse_model3 = map2_dbl(model3, test, ~rmse(model = .x, data = .y))          
  )
```

```{r}
# Visualize cross-validation
cv_results %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(x = "Model",
       y = "RMSE",
       title = "Distribution of root mean squared error in each model")
```

```{r}
# Show average RMSE for each model
cv_results %>% 
  select(starts_with("rmse")) %>% 
  mutate(
    avg_rmse_1 = mean(rmse_model1),
    avg_rmse_2 = mean(rmse_model2),
    avg_rmse_3 = mean(rmse_model3)
  ) %>% 
  select(4:6) %>% 
  distinct() %>% 
  pivot_longer(
    everything(),
    names_to = "Model",
    names_prefix = "avg_rmse_",
    values_to = "Average RMSE"
  ) %>% 
  knitr::kable(digits = 1)
  
```

**Comments:**  
Based on the violin plot and the calculated average root mean squared error (RMSE), the MLR that includes head circumference, length, sex, and all interactions as predictors (model 3) has the lowest average RMSE (288.0).